Distinguishing Spin-Aligned and Isotropic Black Hole Populations With Gravitational Waves

### 2017-04-12 paper/AlignedIsotropicGW.tex
Removed last red.

Figure \ref{fig:O1-posteriors} shows an approximation to the posterior
inferred on $\chieff$ for the three likely \ac{GW} detections
GW150914, GW151226 and LVT151012 from Advanced LIGO's first observing
run (O1) \citep{O1-BBH}.  Because samples drawn from the posterior on
$\chieff$ are not publicly released at this time, we have approximated
the posterior as a Gaussian distribution with the same mean and 90\%
credible interval as quoted in \citet{O1-BBH}.  None of the $\chieff$
posteriors are consistent with two black holes with large aligned
spins, $\chi_{1,2} \gtrsim 0.5$; this contrasts with the large spins
inferred for the majority of black holes in X-ray binaries (see
Section \ref{sec:discussion}).  The analysis here is relatively
insensitive to the precise details of the posterior distributions;
other conclusions are more sensitive.  In particular, our Gaussian
approximation does permit $\chieff = 0$ for GW151226 while the true
posterior rules this out at high confidence
\citep{2016PhRvL.116x1103A,O1-BBH}.

\begin{figure}
  \plotone{../plots/chi-eff-mock-posteriors}
  \caption{\label{fig:O1-posteriors} Approximate posteriors on
    $\chieff$ from the Advanced LIGO O1 run observations in
    \citet{O1-BBH}.  We approximate the posteriors reported in
    \citet{O1-BBH} using Gaussians with the same median and 90\%
    credible interval as reported in \citet{O1-BBH}.  It is notable
    that none of the $\chieff$ posteriors \ilya{supports high \ac{BH}
    spin magnitudes with aligned spins, suggested by observations of stellar-mass black holes in  X-ray binaries} (see
    \citet{2015PhR...548....1M} for a summary of such measurements).}
\end{figure}

@@

Estimates of the rate of \ac{BBH} coalescences give a reasonable
chance of 10 additional \ac{BBH} detections in the next two years
\citep{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming 10
additional detections drawn from each of our six zero-parameter models
for the spin distribution in addition to the three existing detections
from O1, with observational uncertainties drawn randomly from the
three Gaussian widths used to approximate the $\chieff$ posteriors in
Figure \ref{fig:O1-posteriors}\footnote{The measurement uncertainty in
  $\chieff$ depends on the other parameters of the merging \ac{BBH}
  system, particularly on the mass ratio.  Our assumption about future
  observational uncertainties is appropriate if the parameters of the
  existing three events are representative of the parameters of future
  detections.}, we find the odds ratios shown in Figure
\ref{fig:O2-predictions}.  We find that most scenarios with an
additional 10 detections allow the simulated angular distribution to
be inferred with greater than $5\sigma$ ($2.9 \times 10^{-7}$ odds)
credibility; and in the most pessimistic case the distinction is
typically $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$
odds ratio).  While such future detections should permit a confident
distinction between \emph{angular} distributions, we would remain much
less certain about the \emph{magnitude} distribution among the three
options considered here until we have a larger number of observations.
In Figure \ref{fig:O2-predictions}, the odds ratio between different
magnitude distributions with the same angular distribution is much
closer to unity than the odds ratio between angular distributions.

>>>>

Figure \ref{fig:O1-posteriors} shows an approximation to the posterior
inferred on $\chieff$ for the three likely \ac{GW} detections
GW150914, GW151226 and LVT151012 from Advanced LIGO's first observing
run (O1) \citep{O1-BBH}.  Because samples drawn from the posterior on
$\chieff$ are not publicly released at this time, we have approximated
the posterior as a Gaussian distribution with the same mean and 90\%
credible interval as quoted in \citet{O1-BBH}.  None of the $\chieff$
posteriors are consistent with two black holes with large aligned
spins, $\chi_{1,2} \gtrsim 0.5$; this contrasts with the large spins
inferred for the majority of black holes in X-ray binaries with claimed spin measurements (see
Section \ref{sec:discussion}).  The analysis here is relatively
insensitive to the precise details of the posterior distributions;
other conclusions are more sensitive.  In particular, our Gaussian
approximation does permit $\chieff = 0$ for GW151226 while the true
posterior rules this out at high confidence
\citep{2016PhRvL.116x1103A,O1-BBH}.

\begin{figure}
  \plotone{../plots/chi-eff-mock-posteriors}
  \caption{\label{fig:O1-posteriors} Approximate posteriors on
    $\chieff$ from the Advanced LIGO O1 run observations in
    \citet{O1-BBH}.  We approximate the posteriors reported in
    \citet{O1-BBH} using Gaussians with the same median and 90\%
    credible interval as reported in \citet{O1-BBH}.  It is notable
    that none of the $\chieff$ posteriors support high \ac{BH}
    spin magnitudes with aligned spins, suggested by observations of stellar-mass black holes in  X-ray binaries (see
    \citet{2015PhR...548....1M} for a summary of such measurements).}
\end{figure}

@@

Estimates of the rate of \ac{BBH} coalescences give a reasonable
chance of 10 additional \ac{BBH} detections in the next two years
\citep{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming 10
additional detections drawn from each of our six zero-parameter models
for the spin distribution in addition to the three existing detections
from O1, with observational uncertainties drawn randomly from the
three Gaussian widths used to approximate the $\chieff$ posteriors in
Figure \ref{fig:O1-posteriors}\footnote{The measurement uncertainty in
  $\chieff$ depends on the other parameters of the merging \ac{BBH}
  system, particularly on the mass ratio.  Our assumption about future
  observational uncertainties is appropriate if the parameters of the
  three detected events are representative of the parameters of future
  detections.}, we find the odds ratios shown in Figure
\ref{fig:O2-predictions}.  We find that most scenarios with an
additional 10 detections allow the simulated angular distribution to
be inferred with greater than $5\sigma$ ($2.9 \times 10^{-7}$ odds)
credibility; and in the most pessimistic case the distinction is
typically $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$
odds ratio).  While such future detections should permit a confident
distinction between \emph{angular} distributions, we would remain much
less certain about the \emph{magnitude} distribution among the three
options considered here until we have a larger number of observations.
In Figure \ref{fig:O2-predictions}, the odds ratio between different
magnitude distributions with the same angular distribution is much
closer to unity than the odds ratio between angular distributions.
